Automated topic naming to support analysis of software maintenance activities

### 2011-01-14 msr/msr2011.tex
add todo on unsup/sup distinctin
New F1 plot. Some text changes, added a few todos.

\noindent \textbf{Creating a validation corpus} --- For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same NFR labels as \textsf{exp2} (software qualities). 
We looked at each period's topics, and assessed what the data --- consisting of the frequency-weighted word lists and messages --- suggested was the topic for that period. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. 
We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?
%\todo[inline]{inter-rater reliability }
% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}

@@

% word list similarity
Using our three word lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics where there was a match between a word in the list and the same word in the topic.
A \emph{named topic} is a topic with a matching label. 
There are roughly 20 times the number of topics as there are periods per project. 
This is because LDA extracted 20 topics per period. 
All experiments were run on MaxDB 7.500 and MySQL 3.23 data.
Table \ref{tbl:wordlist} shows how many topics were labelled for MaxDB
and MySQL.

@@

\noindent \textbf{Analysis of the unsupervised labelling} -- Based on the labels, and our manual topic labelling, 
we compared the results of the unsupervised word matching approach. 
For each quality we assessed whether unsupervised labels matched the manual annotations.
Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL. 
Our performance is measured using the Receiver Operating Characteristic area-under-curve value~\cite{Fawcett2006861}, abbreviated \emph{ROC}. 
ROC values provide a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. 
ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). 
A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision. 
A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives). 
The area under the ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.

%\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? -- ROC is hot right now and it combines both}
% Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +  it has natural meaningful quality mappings.
\begin{figure*}
  \centering
 \includegraphics[width=0.6\textwidth]{figures/unsupervised-bar}
 \caption{Performance, ROC values (range: 0-1), of unsupervised topic labelling for
   each NFR and per word-list. This graph shows how well the
   unsupervised topic labelling matched our manual annotations.}

%\todo[inline]{this is too tiny in print}
  \label{fig:maxdb-unsup-results}
\end{figure*}

% EXP2
For \textsf{exp2}, the average F-measure (macro-F1) for MaxDB was $0.24$ with a range $0.091$ to
$0.37$, and $0.16$ for MySQL with a range of $0$ to $0.41$.
MaxDB had an average precision and recall of $0.25$ and $0.22$
while MySQL had $0.41$ and $0.10$ respectively.

@@

% \begin{comment}
% maxdb

% portability,0.773778920308,0.228070175439,76,12,0.181684100119,0.52,0.14606741573,0.553033707865,288,0.96,13
% efficiency,0.832904884319,0.21686746988,36,29,0.124642098177,0.236842105263,0.2,0.557848837209,315,0.915697674419,9
% reliability,0.714652956298,0.379888268156,7,104,0.340426350577,0.246376811594,0.829268292683,0.765208858985,244,0.701149425287,34
% functionality,0.902313624679,0.0952380952381,32,6,0.0834236909204,0.25,0.0588235294118,0.520961060481,349,0.983098591549,2
% maintainability,0.796915167095,0.0919540229885,76,3,0.122505166175,0.571428571429,0.05,0.520145631068,306,0.990291262136,4
% usability,0.830334190231,0.175,11,55,0.138109307712,0.112903225806,0.388888888889,0.620320455226,316,0.851752021563,7

% total,0.808483290488,0.235897435897,238,209,0.126930463023,0.248201438849,0.224755700326,0.560823829443,1818,0.896891958559,69mysql

% portability,0.616040955631,0.137931034483,225,0,0.211497874858,1.0,0.0740740740741,0.537037037037,343,1.0,18
% efficiency,0.924914675768,0.3125,33,11,0.297817974114,0.47619047619,0.232558139535,0.606150156324,532,0.979742173112,10
% reliability,0.726962457338,0.411764705882,61,99,0.242477699079,0.361290322581,0.478632478632,0.633772529295,370,0.788912579957,56
% functionality,0.457337883959,0.0245398773006,315,3,0.00597425429824,0.571428571429,0.012539184953,0.500651614948,264,0.988764044944,4
% maintainability,0.686006825939,0,184,0,0,0,0.0,0.5,402,1.0,0
% usability,0.919795221843,0.175438596491,27,20,0.13508225345,0.2,0.15625,0.560074458484,534,0.963898916968,5

% total,0.721843003413,0.159793814433,845,133,0.0857635592199,0.411504424779,0.0991471215352,0.523778370698,2445,0.9484ahindle@wish:~/projects/lda-paper/src/validate$ cat NOTES 
% columns of maxdb.report.text

%     out['accuracy'] = acc
%     out['f1'] = f1
%     out['fn'] = fn
%     out['fp'] = fp
%     out['mcc'] = mcc
%     out['precision'] = pre
%     out['recall'] = rec
%     out['roca'] = roca
%     out['tnr'] = tnr
%     out['tn'] = tn
%     out['tp'] = tp

% cat exp3.*.report.txt 
% maxdb

% portability,0.706940874036,0.472222222222,38,76,0.286382363909,0.40157480315,0.573033707865,0.659850187266,224,0.746666666667,51
% efficiency,0.372750642674,0.222929936306,10,234,0.067549899325,0.130111524164,0.777777777778,0.548772609819,110,0.31976744186,35
% reliability,0.434447300771,0.256756756757,3,217,0.195966959965,0.149019607843,0.926829268293,0.651633024951,131,0.376436781609,38
% functionality,0.53470437018,0.236286919831,6,175,0.186902353269,0.137931034483,0.823529411765,0.665285832643,180,0.507042253521,28
% maintainability,0.627249357326,0.337899543379,43,102,0.111672018872,0.26618705036,0.4625,0.566201456311,207,0.669902912621,37
% usability,0.27763496144,0.107936507937,1,280,0.0937953415493,0.0572390572391,0.944444444444,0.594863731656,91,0.245283018868,17

% total,0.492287917738,0.257983719474,101,1084,0.0926022435775,0.159689922481,0.671009771987,0.568114654124,943,0.46521953626,206
% mysql

% portability,0.616040955631,0.412532637076,164,61,0.170139234681,0.564285714286,0.325102880658,0.57363015753,282,0.822157434402,79
% efficiency,0.532423208191,0.184523809524,12,262,0.124342460899,0.105802047782,0.720930232558,0.619212814253,281,0.517495395948,31
% reliability,0.588737201365,0.404938271605,35,206,0.209194239572,0.284722222222,0.700854700855,0.630811145737,263,0.560767590618,82
% functionality,0.622866894198,0.651968503937,112,109,0.240455709256,0.655063291139,0.648902821317,0.620331560471,158,0.591760299625,207
% maintainability,0.621160409556,0.218309859155,153,69,-0.00390279540262,0.31,0.16847826087,0.498418234912,333,0.828358208955,31
% usability,0.358361774744,0.104761904762,10,366,0.0128981856383,0.0567010309278,0.6875,0.513425090253,188,0.339350180505,22

% total,0.556598407281,0.367032074706,486,1073,0.058598680978,0.296393442623,0.481876332623,0.532831106575,1505,0.583785880528,452
% \end{comment}

>>>>

\noindent \textbf{Creating a validation corpus} --- For MySQL 3.23 and MaxDB 7.500, we manually annotated each extracted topic in each period with the same NFR labels as \textsf{exp2} (software qualities). 
We looked at each period's topics, and assessed what the data --- consisting of the frequency-weighted word lists and messages --- suggested was the topic for that period. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. 
\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?
%\todo[inline]{inter-rater reliability }
% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}

@@

% word list similarity
Using our three word lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics where there was a match between a word in the list and the same word in the topic.
A \emph{named topic} is a topic with a matching label. 
All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA extracted 20 topics per period for each.

Table \ref{tbl:wordlist} shows how many topics were labelled for MaxDB
and MySQL.

@@

\noindent \textbf{Analysis of the unsupervised labelling} -- Based on the labels, and our manual topic labelling, 
we compared the results of the unsupervised word matching approach. 
For each quality we assessed whether unsupervised labels matched the manual annotations.
Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL. 
\todo[inline]{NEIL: this section should be updated to explain F-measure}
Our performance is measured using the Receiver Operating Characteristic area-under-curve value~\cite{Fawcett2006861}, abbreviated \emph{ROC}. 
ROC values provide a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. 
ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). 
A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision. 
A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives). 
The area under the ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.

%\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? -- ROC is hot right now and it combines both}
% Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +  it has natural meaningful quality mappings.
\begin{figure}
  \centering
 \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar}
 \caption{Performance, ROC values (range: 0-1), of unsupervised topic labelling for
   each NFR and per word-list. Dashed line indicates the performance of a random classifier. This graph shows how well the
   unsupervised topic labelling matched our manual annotations.}

%\todo[inline]{this is too tiny in print}
  \label{fig:maxdb-unsup-results}
\end{figure}

\begin{figure}
  \centering
 \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
 \caption{Performance, F1 values (range: 0-1), of unsupervised topic labelling for
   each NFR and per word-list.}
  \label{fig:maxdb-unsup-results-f1}
\end{figure}

\todo[inline]{where is exp1 report?}

% EXP2 source is naming-paper/src/validate/exp2.maxdb.report.txt (change as needed). Second column is F1.
For \textsf{exp2}, the average F-measure (macro-F1) for MaxDB was $0.24$ with a range $0.091$ to
$0.37$, and $0.16$ for MySQL with a range of $0$ to $0.41$.
MaxDB had an average precision and recall of $0.25$ and $0.22$
while MySQL had $0.41$ and $0.10$ respectively.

@@
